D3064R0: How to Avoid OOTA Without Really Trying

### 2024-03-07 ootavacuous.tex
"Non-Trivial Semantic Dependencies": Allow initial-value OOTA
If the initial values happen to match those of an OOTA cycle, then
executions yielding those values might or might not be an OOTA
cycle, depending on whether or not the cycle's loads take values
only from stores within that same cycle.

Suggested-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

In both
Listings~\ref{lst:OOTA Multiplication Example, Initial Value Zero}
and~\ref{lst:OOTA Conditional Multiplication Example, Initial Value Zero},
an OOTA cycle (such as those indicated by their respective \co{exists} clauses)
would need to affect both the \co{x} and the \co{y} components of this
compound semantic dependency.
This situation poses a challenge to current work on semantic dependencies,
which are typically restricted to a single load and store.

@@

Because the only two solutions to $x^2 = x$ are the values zero and one,
Listings~\ref{lst:OOTA Multiplication Example, Initial Value Zero}
and~\ref{lst:OOTA Multiplication Example, Initial Value 1}
each have only one OOTA cycle,\footnote{
	Note well that the initial values cannot be said to have come
	out of thin air.
	Although it is possible to have an OOTA cycle that does not
	involve the initial values, but happens to have the same value,
	the software would not be able to tell the difference.}
and an execution having initial values
of (say) two could have at most two OOTA cycles.
Listing~\ref{lst:OOTA Three-Factor Multiplication Example, Initial Value Zero}
goes one step further by multiplying the values obtained from three
different loads.
Because $x^3 = x$ has three solutions ($\left\{ -1, 0, 1 \right\}$),
this execution could have up to three OOTA cycles.

>>>>

In both
Listings~\ref{lst:OOTA Multiplication Example, Initial Value Zero}
and~\ref{lst:OOTA Conditional Multiplication Example, Initial Value Zero},
an OOTA cycle having non-zero values (such as those indicated by their
respective \co{exists} clauses) would need to affect both the \co{x}
and the \co{y} components of this compound semantic dependency.
This situation poses a challenge to current work on semantic dependencies,
which are typically restricted to a single load and store.

@@

Because the only two solutions to $x^2 = x$ are the values zero and one,
Listings~\ref{lst:OOTA Multiplication Example, Initial Value Zero}
and~\ref{lst:OOTA Multiplication Example, Initial Value 1}
each have only two OOTA cycles.
Note well that although software cannot distinguish the initial
values from those of an OOTA cycle having those same values, there is
a very real difference in the corresponding executions.
To wit, an execution not corresponding to an OOTA cycle will read at
least one initialization value, while an execution corresponding to an
OOTA cycle with those same values will read only from that cycle's stores.
On the other hand, in a similar program having initial values of (say)
two, both executions containing OOTA cycles would be distinguishable
from the initial values.

Listing~\ref{lst:OOTA Three-Factor Multiplication Example, Initial Value Zero}
goes one step further by multiplying the values obtained from three
different loads.
Because $x^3 = x$ has three solutions ($\left\{ -1, 0, 1 \right\}$),
this execution could have up to three OOTA cycles.
